---
title: "Untitled"
author: "Shelly Trigg"
date: "2/22/2019"
output: rmarkdown::github_document
---

```{r}
#Necessary Packages to manipulate data and plot values. 
require(plyr)
require(ggplot2)
require(splitstackshape)
```

```{r}
#Read in  Ct value table
dCt<-read.csv("~/Documents/GitHub/gigas_ploidy_ronits/data/qpcr_ct_values/qpcr_data_consolidated.csv", header=T)

#Split SAMPLE_ID column to create columns for Ploidyulation, treatment, and sample number
dCt<-cSplit(dCt,"Sample", sep= "_", drop=F)

#rename columns appropriately
dCt<-rename(dCt,replace=c("Sample_1"="Ploidy","Sample_2"="Desiccation","Sample_3"="HeatShock","Sample_4"="SampleNum"))

#change NA to 45
dCt[is.na(dCt)] <- 45

#calculate normalized expression of target gene Ct relative to actin Ct using: 2^-(delta Ct)
dCt$HSC70<-2^-(dCt$HSC70-dCt$Actin)
dCt$DNMT1<-2^-(dCt$DNMT1-dCt$Actin)
dCt$MBD2<-2^-(dCt$MBD2-dCt$Actin)
dCt$MeCP2<-2^-(dCt$MeCP2-dCt$Actin)
dCt$HIF1A<-2^-(dCt$HIF1A-dCt$Actin)
dCt$HATHaP2<-2^-(dCt$HATHaP2-dCt$Actin)
dCt$HAT<-2^-(dCt$HAT-dCt$Actin)
dCt$HSP90<-2^-(dCt$HSP90-dCt$Actin)
dCt$SOD<-2^-(dCt$SOD-dCt$Actin)
dCt$ATPsynthetase<-2^-(dCt$ATPsynthetase-dCt$Actin)
dCt$COX1<-2^-(dCt$COX1)



#log transform the data to develop normality in data
dCt$HSC70log<-log(dCt$HSC70)
dCt$DNMT1log<-log(dCt$DNMT1)
dCt$MeCP2log<-log(dCt$MeCP2)
dCt$HIF1Alog<-log(dCt$HIF1A)
dCt$HATHaP2log<-log(dCt$HATHaP2)
dCt$HATlog<-log(dCt$HAT)
dCt$HSP90log<-log(dCt$HSP90)
dCt$MBD2log<-log(dCt$MBD2)
dCt$SODlog<-log(dCt$SOD)
dCt$ATPsynthetaselog<-log(dCt$ATPsynthetase)
dCt$COX1log<-log(dCt$COX1)
```


Run ANOVA's on all log transformed data as well as Tukey's Honestly Significant Difference post hoc test
```{r}
HSC70<-aov(HSC70log~Ploidy+Desiccation+Ploidy:Desiccation, data=dCt)
HSC70
TukeyHSD(HSC70)
summary(HSC70)

DNMT1<-aov(DNMT1log~Ploidy+Desiccation+Ploidy:Desiccation, data=dCt)
DNMT1
TukeyHSD(DNMT1)
summary(DNMT1)

MeCP2<-aov(MeCP2log~Ploidy+Desiccation+Ploidy:Desiccation, data=dCt)
MeCP2
TukeyHSD(MeCP2)
summary(MeCP2)

HIF1A<-aov(HIF1Alog~Ploidy+Desiccation+Ploidy:Desiccation, data=dCt)
HIF1A
TukeyHSD(HIF1A)
summary(HIF1A)

HATHaP2<-aov(HATHaP2log~Ploidy+Desiccation+Ploidy:Desiccation, data=dCt)
HATHaP2
TukeyHSD(HATHaP2)
summary(HATHaP2)

HAT<-aov(HATlog~Ploidy+Desiccation+Ploidy:Desiccation, data=dCt)
HAT
TukeyHSD(HAT)
summary(HAT)

HSP90<-aov(HSP90log~Ploidy+Desiccation+Ploidy:Desiccation, data=dCt)
HSP90
TukeyHSD(HSP90)
summary(HSP90)

SOD<-aov(SODlog~Ploidy+Desiccation+Ploidy:Desiccation, data=dCt)
SOD
TukeyHSD(SOD)
summary(SOD)

ATPsynthetase<-aov(ATPsynthetaselog~Ploidy+Desiccation+Ploidy:Desiccation, data=dCt)
ATPsynthetase
TukeyHSD(ATPsynthetase)
summary(ATPsynthetase)

MBD2<-aov(MBD2log~Ploidy+Desiccation+Ploidy:Desiccation, data=dCt)
MBD2
TukeyHSD(MBD2)
summary(MBD2)

COX1<-aov(COX1log~Ploidy+Desiccation+Ploidy:Desiccation, data=dCt)
COX1
TukeyHSD(COX1)
summary(COX1)
```

graph all normalized Ct values to produce boxplots to visualize data
```{r, echo = FALSE, fig.height = 12, fig.width = 15}

ggplot(data=dCt)+geom_boxplot(aes(x=Ploidy, y=HSC70,fill=Desiccation))+theme_bw()+
  scale_fill_manual(values=c("#31C4ED", "#62D12B"),
                  labels=c("Control","Desiccation + Elevated Temp."))+
  guides(fill=guide_legend(title="Treatment"))+
  theme(axis.text.x=element_text(size=15), axis.text.y=element_text(size=13),
        axis.title.x=element_text(size=19), axis.title.y=element_text(size=20),
        legend.position=c(.09,.87),panel.grid.major=element_blank(),
        legend.key=element_rect(fill=NA))+
  ylim(c(0,0.023))+scale_x_discrete(labels=c("Diploid","Triploid"))+
  labs(x="Ploidy", y=expression(paste("HSC70 Expression (",Delta,"Ct)")))

ggplot(data=dCt)+geom_boxplot(aes(x=Desiccation, y=DNMT1,fill=Ploidy))+theme_bw()+
  scale_fill_manual(values=c("#31C4ED", "#62D12B"),
                  labels=c("Diploid","Triploid"))+
  guides(fill=guide_legend(title="Ploidy"))+
  theme(axis.text.x=element_text(size=15), axis.text.y=element_text(size=13),
        axis.title.x=element_text(size=19), axis.title.y=element_text(size=20),
        legend.position=c(.09,.87),panel.grid.major=element_blank(),
        legend.key=element_rect(fill=NA))+
  ylim(c(0,0.00075))+scale_x_discrete(labels=c("Control","Desiccation + Elevated Temp."))+
  labs(x="Treatment", y=expression(paste("DNMT1 Expression (",Delta,"Ct)")))

ggplot(data=dCt)+geom_boxplot(aes(x=Desiccation, y=MBD2,fill=Ploidy))+theme_bw()+
  scale_fill_manual(values=c("#31C4ED", "#62D12B"),
                  labels=c("Diploid","Triploid"))+
  guides(fill=guide_legend(title="Ploidy"))+
  theme(axis.text.x=element_text(size=15), axis.text.y=element_text(size=13),
        axis.title.x=element_text(size=19), axis.title.y=element_text(size=20),
        legend.position=c(.09,.87),panel.grid.major=element_blank(),
        legend.key=element_rect(fill=NA))+
  ylim(c(0,0.0002))+scale_x_discrete(labels=c("Control","Desiccation + Elevated Temp."))+
  labs(x="Treatment", y=expression(paste("MBD2 Expression (",Delta,"Ct)")))

ggplot(data=dCt)+geom_boxplot(aes(x=Desiccation, y=MeCP2,fill=Ploidy))+theme_bw()+
  scale_fill_manual(values=c("#31C4ED", "#62D12B"),
                  labels=c("Diploid","Triploid"))+
  guides(fill=guide_legend(title="Ploidy"))+
  theme(axis.text.x=element_text(size=15), axis.text.y=element_text(size=13),
        axis.title.x=element_text(size=19), axis.title.y=element_text(size=20),
        legend.position=c(.09,.87),panel.grid.major=element_blank(),
        legend.key=element_rect(fill=NA))+
  ylim(c(0,0.0004))+scale_x_discrete(labels=c("Control","Desiccation + Elevated Temp."))+
  labs(x="Treatment", y=expression(paste("MeCP2 Expression (",Delta,"Ct)")))

ggplot(data=dCt)+geom_boxplot(aes(x=Desiccation, y=HIF1A,fill=Ploidy))+theme_bw()+
  scale_fill_manual(values=c("#31C4ED", "#62D12B"),
                  labels=c("Diploid","Triploid"))+
  guides(fill=guide_legend(title="Ploidy"))+
  theme(axis.text.x=element_text(size=15), axis.text.y=element_text(size=13),
        axis.title.x=element_text(size=19), axis.title.y=element_text(size=20),
        legend.position=c(.09,.87),panel.grid.major=element_blank(),
        legend.key=element_rect(fill=NA))+
  ylim(c(0,0.0004))+scale_x_discrete(labels=c("Control","Desiccation + Elevated Temp."))+
  labs(x="Treatment", y=expression(paste("HIF1A Expression (",Delta,"Ct)")))

ggplot(data=dCt)+geom_boxplot(aes(x=Desiccation, y=HATHaP2,fill=Ploidy))+theme_bw()+
  scale_fill_manual(values=c("#31C4ED", "#62D12B"),
                  labels=c("Diploid","Triploid"))+
  guides(fill=guide_legend(title="Ploidy"))+
  theme(axis.text.x=element_text(size=15), axis.text.y=element_text(size=13),
        axis.title.x=element_text(size=19), axis.title.y=element_text(size=20),
        legend.position=c(.09,.87),panel.grid.major=element_blank(),
        legend.key=element_rect(fill=NA))+
  ylim(c(0,0.013))+scale_x_discrete(labels=c("Control","Desiccation + Elevated Temp."))+
  labs(x="Treatment", y=expression(paste("HATHaP2 Expression (",Delta,"Ct)")))

ggplot(data=dCt)+geom_boxplot(aes(x=Desiccation, y=HAT,fill=Ploidy))+theme_bw()+
  scale_fill_manual(values=c("#31C4ED", "#62D12B"),
                  labels=c("Diploid","Triploid"))+
  guides(fill=guide_legend(title="Ploidy"))+
  theme(axis.text.x=element_text(size=15), axis.text.y=element_text(size=13),
        axis.title.x=element_text(size=19), axis.title.y=element_text(size=20),
        legend.position=c(.09,.87),panel.grid.major=element_blank(),
        legend.key=element_rect(fill=NA))+
  ylim(c(0,0.0003))+scale_x_discrete(labels=c("Control","Desiccation + Elevated Temp."))+
  labs(x="Treatment", y=expression(paste("HAT Expression (",Delta,"Ct)")))

ggplot(data=dCt)+geom_boxplot(aes(x=Desiccation, y=HSP90,fill=Ploidy))+theme_bw()+
  scale_fill_manual(values=c("#31C4ED", "#62D12B"),
                  labels=c("Diploid","Triploid"))+
  guides(fill=guide_legend(title="Ploidy"))+
  theme(axis.text.x=element_text(size=15), axis.text.y=element_text(size=13),
        axis.title.x=element_text(size=19), axis.title.y=element_text(size=20),
        legend.position=c(.09,.87),panel.grid.major=element_blank(),
        legend.key=element_rect(fill = NA))+
  ylim(c(0,0.013))+scale_x_discrete(labels=c("Control","Desiccation + Elevated Temp."))+
  labs(x="Treatment", y=expression(paste("HSP90 Expression (",Delta,"Ct)")))

ggplot(data=dCt)+geom_boxplot(aes(x=Desiccation, y=SOD,fill=Ploidy))+theme_bw()+
  scale_fill_manual(values=c("#31C4ED", "#62D12B"),
                  labels=c("Diploid","Triploid"))+
  guides(fill=guide_legend(title="Ploidy"))+
  theme(axis.text.x=element_text(size=15), axis.text.y=element_text(size=13),
        axis.title.x=element_text(size=19), axis.title.y=element_text(size=20),
        legend.position=c(.09,.87),panel.grid.major=element_blank(),
        legend.key=element_rect(fill = NA))+
  ylim(c(0,0.4))+scale_x_discrete(labels=c("Control","Desiccation + Elevated Temp."))+
  labs(x="Treatment", y=expression(paste("SOD Expression (",Delta,"Ct)")))

ggplot(data=dCt)+geom_boxplot(aes(x=Desiccation, y=ATPsynthetase,fill=Ploidy))+theme_bw()+
  scale_fill_manual(values=c("#31C4ED", "#62D12B"),
                  labels=c("Diploid","Triploid"))+
  guides(fill=guide_legend(title="Ploidy"))+
  theme(axis.text.x=element_text(size=15), axis.text.y=element_text(size=13),
        axis.title.x=element_text(size=19), axis.title.y=element_text(size=18),
        legend.position=c(.09,.87),panel.grid.major=element_blank(),
        legend.key=element_rect(fill = NA))+
  ylim(c(0,0.0016))+scale_x_discrete(labels=c("Control","Desiccation + Elevated Temp."))+
  labs(x="Treatment", y=expression(paste("ATP Synthetase Expression (",Delta,"Ct)")))

ggplot(data=dCt)+geom_boxplot(aes(x=Desiccation, y=COX1,fill=Ploidy))+theme_bw()+ 
  scale_fill_manual(values=c("#31C4ED", "#62D12B"), 
                    labels=c("Diploid","Triploid")) + 
  guides(fill=guide_legend(title="Ploidy")) +
  theme(axis.text.x=element_text(size=15), axis.text.y=element_text(size=13),
        axis.title.x=element_text(size=19), axis.title.y=element_text(size=18),
        legend.position=c(.09,.87),panel.grid.major=element_blank(),
        legend.key=element_rect(fill = NA))+scale_x_discrete(labels=c("Control","Desiccation + Elevated Temp."))+ labs(x="Treatment", y=expression(paste("COX1 DNA Abundance")))