---
title: "GlobalDNAMeth_Polyploids"
author: "Shelly Trigg"
date: "2/12/2019"
output: rmarkdown::github_document
---

Load libraries
```{r}
library(readxl)
library(dplyr)
library(ggplot2)
library(broom)
```

Read in raw absorbance data files exported from plate reader. There are 3 files because I read the plate 3 independent times. The purpose for this was to make sure the readings were stable. 
```{r}
data1 <- read_xls("~/Documents/GitHub/C_gigas/Polyploids/docs/20190206_FirstReadingSTRIGG.xls")
data2 <- read_xls("~/Documents/GitHub/C_gigas/Polyploids/docs/20190206_2ndReadingSTRIGG.xls")
data3 <- read_xls("~/Documents/GitHub/C_gigas/Polyploids/docs/20190206_3rdReadingSTRIGG.xls")
```

Combine data frames into one data frame
```{r}
Feb2019 <- merge(data1[,c(3,6)],data2[,c(3,6)], by = "Well")
Feb2019 <- merge(Feb2019,data3[,c(3,6)], by = "Well")
#simplify column names
colnames(Feb2019) <- c("Well", "Abs1", "Abs2", "Abs3")
```

Check that absorbance readings were stable
```{r}
#calculate the mean reading for each of the three readings
Feb2019$mean <- apply(Feb2019[,2:4],1, mean)
#calculate the standard deviation for each of the three readings
Feb2019$sd <- apply(Feb2019[,2:4],1, sd)
#print the largest standard deviation
max(Feb2019$sd)
```

**Stable readings conclusion:**  SDs are very small between readings so go with readings from data1

Add sample names to data and remove wells that shouldn't be considered
```{r}
#first read in sample names from plate map list
Feb_plateMap <- read_xlsx("~/Documents/GitHub/C_gigas/Polyploids/docs/20190131_Ronit'sSamplesforGlobalDNAMeth.xlsx", sheet = 3)

#make a list of samples to not be considered
rmv_wells <- c("no_sample", "contam", "no_ab")

#remove samples not to be considered from data frame
Feb_data <- merge(Feb_plateMap[which(!(Feb_plateMap$Sample %in% rmv_wells)),], data1[,c(3,6)])

#simplify absorbance column name
colnames(Feb_data)[4] <- "Absorbance"
```


Calculate average absorbance and SD for duplicate samples
```{r}
#make a dataframe with mean absorbances for duplicate samples
Feb_data_avg <- Feb_data %>% group_by(Sample) %>% summarize(avgAbs=mean(Absorbance))
#make a dataframe with absorbance standard deviations for duplicate samples
Feb_data_sd <- Feb_data  %>% group_by(Sample) %>% summarize(SD=sd(Absorbance))
#merge mean absorbance and standard deviation dataframes into one dataframe
Feb_data_avg <- merge(Feb_data_avg,Feb_data_sd, by = "Sample")
#merge mean and standard deviation dataframe with sample info
Feb_data_avg <- unique(merge(Feb_data[,c("Type","Sample")], Feb_data_avg, by = "Sample"))
```

Generate standard curve
```{r}
#make new data frame with only data from standards (STDs)
curve <- Feb_data_avg[grep("STD", Feb_data_avg$Type),]
print(curve)

#make new column for %meth
curve$perc_meth <- c(0.1,0.2,0.5,1,5,0,10)
#order data by perc_meth
curve <- curve[order(curve$perc_meth),]
print(curve)
```

plot curve
```{r}
ggplot(curve, aes(perc_meth, avgAbs)) + geom_point() 
```

last two points are plateauing so remove them
```{r}
ggplot(curve[which(curve$perc_meth < 5.0),], aes(perc_meth, avgAbs)) + geom_point() 
```

find regression line
```{r}
fit <- lm(curve[which(curve$perc_meth < 5.0),"avgAbs"] ~ curve[which(curve$perc_meth < 5.0),"perc_meth"] )
#find the R-squared
rsquared <- summary(fit)$r.squared
rsquared
#[1] 0.9978712

#find the slope (https://www.cyclismo.org/tutorial/R/linearLeastSquares.html)
slope <- fit$coefficients[[2]]
slope
```

plot regression line with equation and r-squared 
```{r}
#found this site helpful: (https://www.listendata.com/2016/07/add-linear-regression-equation-and.html)
linear = function(k) {
  z <- list(xx = format(coef(k)[1], digits = 5),
            yy = format(abs(coef(k)[2]), digits = 5),
            r2 = format(summary(k)$r.squared, digits = 5));
  if (coef(k)[1] >= 0)  {
    eq <- substitute(italic(Y) == yy %*% italic(X) + xx*","~~italic(r)^2~"="~r2,z)
  } else {
    eq <- substitute(italic(Y) == yy %*% italic(X) - xx*","~~italic(r)^2~"="~r2,z)  
  }
  as.character(as.expression(eq));              
}

x <- curve[which(curve$perc_meth < 5.0),"perc_meth"]
y <- curve[which(curve$perc_meth < 5.0),"avgAbs"]
fo = y ~ x
linplot <- ggplot(data = curve[which(curve$perc_meth < 5.0),], aes(x = perc_meth, y = avgAbs)) + geom_smooth(method = "lm", se=FALSE, color="black", formula = fo) +  geom_point() + theme_bw()
linplot1 = linplot + annotate("text", x = 0.3, y = 0.7, label = linear(lm(fo)), colour="black", size = 5, parse=TRUE)
linplot1
```

Calculate perc_meth for samples using equation in EpiGentek MethylFlash kit manual:
5-mC% = (sample OD - NC OD) / (slope x input sample DNA in ng) * 100%

```{r}
#create an object for the absorbance of the NC (blank)
NC_OD <- Feb_data_avg[grep("5 ul NC", Feb_data_avg$Sample),"avgAbs"]
# create an object for product of the regression line slope and the ng DNA added in each of the wells
slopexDNAamt <- slope * 25

#create a new column in the dataframe with the NC_OD subtracted from the avg absorbance
Feb_data_avg$avgAbs_NC <- Feb_data_avg$avgAbs - NC_OD
#create a new column with the calculated percent methylation
Feb_data_avg$perc_meth <- Feb_data_avg$avgAbs_NC/slopexDNAamt *100
```

Add ploidy and treatment info to data
```{r}
#make new data frame with only sample info
Feb_data_avg_Samples <- Feb_data_avg[which(Feb_data_avg$Type == "Sample"),]

#Add sample description to data frame. This loops through each line in Feb_data_avg_Samples and creates a new column 'ploidy' that gets filled in with "diploid" or "triploid" depending on whether the "Sample" name starts with a "D" or "T"; and creates a new column "treatment" that gets filled in with "heat_stress" or "control" depending on whether the number following the D or T is less than 10 or greater than 10. 
for (i in 1:nrow(Feb_data_avg_Samples)){
  if(substr(Feb_data_avg_Samples$Sample[i],1,1)=="D"){
    Feb_data_avg_Samples$ploidy[i] <- "diploid"
  }
  if(substr(Feb_data_avg_Samples$Sample[i],1,1)=="T"){
    Feb_data_avg_Samples$ploidy[i] <- "triploid"
  }
  if(as.numeric(substr(Feb_data_avg_Samples$Sample[i],2,3)) > 10){
    Feb_data_avg_Samples$treatment[i] <- "heat_stress"
  }
  if(as.numeric(substr(Feb_data_avg_Samples$Sample[i],2,3)) < 10){
    Feb_data_avg_Samples$treatment[i] <- "control"
  }
}

#make a new column which contains both ploidy and treatment info 
Feb_data_avg_Samples$condition <- paste(Feb_data_avg_Samples$ploidy, Feb_data_avg_Samples$treatment, sep = "_")
```

Plot absorbances vs. experimental group
```{r}
ggplot(Feb_data_avg_Samples, aes(x = condition, y = perc_meth)) + geom_boxplot(aes(fill = condition), show.legend = F) + scale_fill_manual(values = c("dodgerblue4", "lightblue1", "darkgreen", "palegreen")) + xlab("Sample") + theme_bw() + ylab("% 5-mC") + ggtitle("Ploidy x heat stress effect on global 5-mC DNA methylation")
```

Perform an ANOVA test to see if ploidy, treatment, or their interaction has a signficant effect on % methylation at p < 0.05. 
```{r}
#run anova
aov_2way <- aov(perc_meth ~ ploidy + treatment + ploidy:treatment, data = Feb_data_avg_Samples)
#save the model summary as an object
aov_2way_model_summary <- glance(aov_2way)
#print the p.value
aov_2way_model_summary$p.value
```

The p-value is significant at 0.05 so run a Tukey's HSD test to see which effect is significant
```{r}
tuk <- TukeyHSD(aov(perc_meth ~ ploidy + treatment + ploidy:treatment, data = Feb_data_avg_Samples))
tuk
```

**Conclusions: Effect of ploidy and heat stress on global 5-mC DNA methylation**  
Based on the Tukey's test results:  

1. The diploid heat stress %5-mC methylation is significantly different from 
	- the diploid control 
 	- the triploid control 
 	- the triploid heatstress 
2. The triploid heat stress %5-mC is not different than the diploid or triploid controls. 

In summary, diploid ctenidia shows a significant increase in global 5-mC methylation in response to heatstress while triploid ctenidia does not. 

